Inference of proto-neutron star properties from gravitational-wave data in core-collapse supernovae.

### 2020-10-29 introduction.tex
A little more work on the intro ...

In the last decade, significant progress has been made in the development of numerical codes, {in particular in the treatment of multidimensioal effects \citep{BMueller:2020}.}
% {[PCD: I don't think the next sentence is necessary since we have already discarded magneto-rotational events for this work.]
%\sout{The waveforms produced by the magneto-rotational mechanism in particular is well understood. The core-bounce signal can be directly related with the rotational properties of the core} \citep{Dimmelmeier:2007, abdikamalov:14, Richers:2017}. \sout{However, the low rate of this kind of events and its expected low amplitude in the slow-rotation case, will probably impede its detection.}}
In the case of  neutrino-driven explosions, the GW emission is {primarly induced by instabilities developed at the newly formed proto-neutron star (PNS) and by the non-spherical accreting flow of hot matter over its surface \citep{Kotake:2017}.  This} dynamics excite the different modes of oscillation of the PNS{, which ultimately leads to the emission of GWs. The frequency and time evolution of these modes carry information about the
properties of the GW emitter and could allow to perform PNS asteroseismology. }

@@

{The main feature appearing systematically in the GW spectrum of multidimensional numerical simulations is a strong and relatively narrow \mab{oscillation} in the post bounce evolution with raising frequency
from about $100$~Hz up to a few kHz (at most) and a typical duration of $0.5-1$~s. This feature has been interpreted as a continuously excited gravity mode (g-mode, see \citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS \citep{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015, Kuroda:2016, Andresen:2017}. In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS.} The {typical} frequencies of {these} modes make them {a} promising {source} for ground-based interferometers.

 The {properties of} g-modes in hot {PNSs} have been studied since the end of last century {by means of linear perturbation analysis of background PNS models}. The oscillation modes {associated to} the surface of hot PNSs were first considered by McDermott {\it et al.} \citep{McDermott:1983}. Additionally, the stratified structure of the PNS allows for the presence of different types of g-modes related to the fluid core \citep{Reisenegger:1992}. Many posterior works used simplified neutron star models assuming an equilibrium configuration {as a background}, to study the effect of rotation \citep{Ferrari:2004}, general relativity \citep{Passamonti:2005}, non-linearities \citep{Dimmelmeier:2006}, phase transition \citep{Kruger:2015} and realistic equation of state \citep{Camelio:2017}. {Only recently, there has been an effort to incorporate realistic backgrounds based in numerical simulations in the computation of the mode structure and evolution \citep{Sotani:2016,Torres:2018, Morozova:2018, Torres:2019a,Torres:2019b,Sotani:2019,WS:2019,Sotani:2020a, Sotani:2020b}}.

>>>>

In the last decade significant progress has been made in the development of numerical codes, {in particular in the treatment of multidimensioal effects \citep{BMueller:2020}.} In the case of  neutrino-driven explosions, the GW emission is {primarly induced by instabilities developed at the newly formed proto-neutron star (PNS) and by the non-spherical accreting flow of hot matter over its surface \citep{Kotake:2017}.  This} dynamics excite the different modes of oscillation of the PNS which ultimately leads to the emission of GWs. The frequency and time evolution of these modes carry information about the properties of the GW emitter and could allow to perform PNS asteroseismology.

@@

All multidimensional numerical simulations show the systematic appearance in time-frequency diagrams (or spectrograms) of a distinct and relatively narrow feature 
%
%{The main feature appearing systematically in the GW spectrum of multidimensional numerical simulations is a strong and relatively narrow \mab{oscillation} 
%
during the post-bounce evolution of the system, with raising frequency
from about $100$~Hz up to a few kHz (at most) and a typical duration of $0.5-1$~s. This feature has been interpreted as a continuously excited gravity mode (g-mode, see \citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS \citep{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015, Kuroda:2016, Andresen:2017}. 
In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS. The {typical} frequencies of {these} modes make them interesting targets for
%promising {source} for 
ground-based GW interferometers.

 The {properties of} g-modes in hot {PNSs} have been studied since the 1990s
 %end of last century 
 {by means of linear perturbation analysis of background PNS models}. The oscillation modes connected with the surface of hot PNSs were first considered by McDermott {\it et al.} \citep{McDermott:1983}. Additionally, the stratified structure of the PNS allows for the presence of different types of g-modes related to the fluid core \citep{Reisenegger:1992}. Many subsequent works used simplified neutron star models assuming an equilibrium configuration {as a background}, to study the effect of rotation \citep{Ferrari:2004}, general relativity \citep{Passamonti:2005}, non-linearities \citep{Dimmelmeier:2006}, phase transitions \citep{Kruger:2015} and realistic equation of state \citep{Camelio:2017}. {Only recently, there have been efforts to incorporate more suitable backgrounds based on numerical simulations in the computation of the mode structure and evolution \citep{Sotani:2016,Torres:2018, Morozova:2018, Torres:2019a,Torres:2019b,Sotani:2019,WS:2019,Sotani:2020a, Sotani:2020b}}.
